### 2025-01-23 src/tex/ms.tex
[showyourwork] overleaf sync

%                                                                 aa.dem
% AA vers. 9.1, LaTeX class for Astronomy & Astrophysics
% demonstration file
%                                                       (c) EDP Sciences
%-----------------------------------------------------------------------
%
%\documentclass[referee]{aa} % for a referee version
%\documentclass[onecolumn]{aa} % for a paper on 1 column  
%\documentclass[longauth]{aa} % for the long lists of affiliations 
%\documentclass[letter]{aa} % for the letters 
%\documentclass[bibyear]{aa} % if the references are not structured 
%                              according to the author-year natbib style

%
\documentclass{aa}

%
\usepackage{showyourwork}

\usepackage{graphicx}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{txfonts}
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{siunitx}
\usepackage{hyperref}
\usepackage{url}
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{booktabs}       % professional-quality tables
\usepackage{tabularx}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\usepackage[options]{hyperref}
% To add links in your PDF file, use the package "hyperref"
% with options according to your LaTeX or PDFLaTeX drivers.
%

\usepackage{tikz}

\usepackage{xcolor}

\newcommand{\todo}[1]{\textcolor{red}{#1}}
\newcommand{\T}[1]{\textcolor{orange}{#1}}
%\newcommand{\U}[1]{\textbf{\textit{\textcolor{blue}{#1}}}}
\newcommand{\U}[1]{\textcolor{black}{#1}}
\usepackage{subfigure}

\begin{document}

   \title{Inferring Galactic Parameters from Chemical Abundances with Simulation-Based Inference}

   \titlerunning{Simulation-Based Inference for Galactic Parameters}
   \authorrunning{T. Buck and B. Günes}

   %\subtitle{Inferring Galactic Parameters from Chemical Abundances with Simulation-Based Inference}

   \author{Tobias Buck\inst{1,2}
          \and
          Berkay Günes\inst{1,2}
          }

   \institute{Interdisciplinary Center for Scientific Computing (IWR), University of Heidelberg,
 Im Neuenheimer Feld 205, D-69120 Heidelberg
 \and
 Universität Heidelberg, Zentrum für Astronomie, Institut für Theoretische Astrophysik, Albert-Ueberle-Straße 2, D-69120 Heidelberg, Germany\\
 \email{tobias.buck@iwr.uni-heidelberg.de}
             }

   \date{Received Month, XXXX; accepted Month Day, XXXX}

% \abstract{}{}{}{}{} 
% 5 {} token are mandatory
 \abstract
  % context heading (optional)
  % {} leave it empty if necessary  
   {Chemical abundances of stars are able to reveal important galactic parameters, such as the initial mass function high-mass slope and the frequency of type Ia supernovae. Constraining these parameters is of critical importance in order to create realistic hydrodynamical simulations and understand galactic physics. However, so far inference was significantly limited by the computational cost of traditional Bayesian algorithms. Here we present a new method using simulation-based inference (SBI) to circumvent previous limitations and enable inference from massive stellar surveys.
    SBI enables a new approach to the inference of the posterior distribution of Galactic parameters under an intractable likelihood function by utilizing forward simulations of stellar abundance distributions.
    Using SBI we are able to predict global galactic parameters of great accuracy and precision from chemical abundances of multiple stars, faster than conventional methods like Hamiltonian Monte Carlo inference. Our method easily scales to hundred-thousands of stars and beats previous approaches by orders of magnitude in compute time and accuracy.}
  % aims heading (mandatory)
   {}
  % methods heading (mandatory)
   {}
  % results heading (mandatory)
   {}
  % conclusions heading (optional), leave it empty if necessary 
   {}

   \keywords{Galaxies: fundamental parameters --
            Galaxies: stellar content --
             Methods: data analysis --
             Methods: statistical --
             }
\maketitle
%-------------------------------------------------------------------
\section{Introduction}

Currently all cosmological simulations of galaxy formation \citep[e.g.][]{Sawala2016,Hopkins2018,Pillepich2018,Buck2020,Buck2020c,Font2020,Agertz2021} rely on a small set of galactic parameters to describe fundamental effects of stellar evolution, including the birth and death rates for various types of stars or the production of heavy elements inside stars \citep[e.g.][]{Buck2021}. Two crucial unknowns are the shape of the initial mass function (IMF), setting the mass distribution of stars born from the interstellar medium (ISM), and the rate of Type Ia supernovae (SN\,Ia) explosions.

The exact values of parameters sensitively affect chemical evolution tracks \citep{2005A&A...430..491R,2015MNRAS.449.1327V,2015MNRAS.451.3693M}, yet their observational constraints remain weak.
%
For example, a range of high-mass IMF slopes have been suggested \citep[Tab.\,7]{2016ApJ...824...82C}, with a steeper-than-canonical slope being suggested by a range of studies \citep[e.g.][]{2015ApJ...806..198W,Rybizki2015,Chabrier2014}. In addition, the IMF slope may itself be not a constant but rather a function of metallicity, introducing further complexity \citep[e.g.][]{2019MNRAS.482..118G,Martin2019}. Similarly, the choice of SN\,Ia delay-time-distribution and normalization plays a crucial role in the enrichment of the ISM \citep[e.g.][]{Buck2021} and is heavily debated \citep{2010ApJ...722.1879M,2012MNRAS.426.3282M,2015ApJ...810..137J}.

The goal of this work is to demonstrate how we can use modern machine learning (ML) techniques in tandem with a simple galactic chemical evolution (GCE) model in a Bayesian framework to infer global galactic parameters from a set of stars. Our work focuses on two key parameters; the high-mass slope of the \citet[Tab.\,1]{2003PASP..115..763C} IMF and the rate of SN\,Ia explosions per unit mass. We assume that both are constant across the galaxy and time.

A methodologically very similar work was previously presented by \cite{Philcox_2019} using Hamilton Monte Carlo sampling for the inference. The result of this work was an impressive speed-up over classical Markov Chain Monte Carlo sampling with no loss of accuracy. Here we make use of simulation-based inference \citep[SBI, e.g.][]{Cranmer2020}. For high-dimensional posterior functions, SBI gives tremendous computational improvements over more traditional methods such as Markov Chain Monte Carlo (MCMC) or even the more efficient Hamilton Monte Carlo (HMC) methods.

This paper is structured as follows. In Section~\ref{sec:methods} we detail our methods. In particular, we discuss the GCE model and the methods behind SBI. Our results are presented in Section~\ref{sec: Results} where we test the performance of our approach on mock data and data taken from a hydrodynamical simulation. We conclude our study with Section~\ref{sec: conclusion} with a summary and outlook of our results.

\textcolor{red}{to be adapted... especially the zenodo link\\
Finally, we publicly release all of our code to reproduce the results of this manuscript via GitHub\footnote{URL: {\url{https://github.com/TobiBu/sbi-chempy}}}and refer to Appendix \ref{sec:appendix_code_and_data} for an overview of our code and structure. All our datasets and network weights are publicly available on Zenodo.\footnote{URL: \url{https://zenodo.org/record/8375344}}}

\begin{figure*}[]
     \centering
     %\vspace{-.25cm}
     \includegraphics[width=1\linewidth]{figures/sbi_overview.png}
     \vspace{-.5cm}
     \caption{SBI flow chart. From a set of priors we simulate a sample of stellar abundances using \texttt{CHEMPY} \citep{Rybizki_2017,Philcox_2019} which we use to train a \emph{neural network} emulator to speed up the data generation process. Using the \emph{neural network} emulator we produce training data to train the Neural Density Estimator. With this we infer the posterior distribution of the model parameters from a single star. Repeating that for $N_{\rm stars}$ from the same galaxy gives an accurate fit of the IMF slope and Type Ia supernovae normalization.}
     \label{fig:flowchart}
\end{figure*}

% %--------------------------------------------------------------------
\section{Methods}
\label{sec:methods}

In order to establish our new method based on simulation based inference we need two ingredients: A simulator (in our case a Galactic chemical evolution model) that simulates observational data from a set of model parameters (in our case the IMF slope and the Type Ia supernovae normalization) and a flexible way of parametrizing the posterior density conditioned on the observation in order to perform our inference. In the next subsections we describe both ingredients in detail.

\subsection{Galactic chemical evolution models}
Our simulator is based on the \texttt{CHEMPY} model \citep{Rybizki_2017}. \texttt{CHEMPY} is a simple Galactic Chemical Evolution (GCE) model that is able to predict stellar chemical abundances throughout cosmic time by using published nucleosynthetic yield tables for three key processes (SN\,Ia and SN\,II explosions and AGB stellar feedback) and a small number of parameters controlling simple stellar populations (SSPs) and ISM physics. We refer to the initial \texttt{CHEMPY} paper \citep{Rybizki_2017} for the details of the model.

In particular, we are using the \texttt{CHEMPYScoring} module \citep{Philcox_2018} publicly available as the \texttt{CHEMPYMulti} \citep{Philcox_2019}\footnote{\href{https://github.com/oliverphilcox/ChempyMulti}{github.com/oliverphilcox/ChempyMulti}} package a further development of the original \texttt{CHEMPY} model.

\paragraph{\texttt{CHEMPY} parameters}
In this work, we allow six \texttt{CHEMPY} parameters to vary freely (see also Tab.\,\ref{tab:priors}). These can be categorized into three groups:
\begin{enumerate}
     \item $\vec\Lambda$: \textbf{Global Galactic Parameters} describe SSP physics and comprise the high-mass \citet{2003PASP..115..763C} IMF slope, $\alpha_\mathrm{IMF}$, and (logarithmic) Type Ia SN normalization, $\log_{10}(N_\mathrm{Ia})$. We treat these as star-independent and assume them to be constant across galactic environments and cosmic time\footnote{Whilst $\log_{10}(N_\mathrm{Ia})$ is constant with respect to time by definition, it being simply a normalization constant, there is some evidence for $\alpha_\mathrm{IMF}$ varying as a function of time or metallicity \citep{Chabrier2014,2016MNRAS.462.2832C,2019MNRAS.482..118G,Martin2019}.}. 
     We adopt the same broad priors as \citep{Philcox_2019} for these variables (see also Tab.\ref{tab:priors}). 
     %
     \item $\{\vec\Theta_i\}$: \textbf{Local Galactic Parameters} describe the local physics of the ISM and are hence specific to each stellar environment, indexed by $i$. As defined in \citep{Rybizki_2017}, these include the star-formation efficiency (SFE), $\log_{10}(\text{SFE})$, $\log_{10}(\mathrm{SFR}_\mathrm{peak})$, which controls the peak of the star formation rate (SFR), and the fraction of stellar outflow that is fed to the gas reservoir, $\mathrm{x}_\mathrm{out}$. We adopt broad priors for all parameters and, as in \citep{Philcox_2019}, fix the SN\,Ia delay-time distribution, $\log_{10}(\tau_{\rm Ia})$, to $\log_{10}(\tau_{\rm Ia})=-0.80$ \citep[see also][]{Philcox_2018}.
     \item $\{T_i\}$: \textbf{Stellar Birth-Times}. Time in Gyr at which a given star is formed from the ISM. We assume that its proto-stellar abundances match the local ISM abundances at $T_i$.
\end{enumerate}

The separability of local (ISM) parameters and global (SSP) parameters is motivated by recent observational evidence: \citet{2019arXiv190710606N} find that the elemental abundances of red clump stars belonging to the thin disk can be predicted almost perfectly from their age and [Fe/H] abundance. This implies that the key chemical evolution parameters affecting the elemental abundances (SSP parameters and yield tables) are held fixed, whilst ISM parameters vary smoothly over the thin disk \citep[which offsets the metallicity for different galactocentric radii, e.g.][for a simulated example]{Buck2020, Wang2024}. Similarly \cite{2019ApJ...874..102W} find that ISM parameter variations are deprojected in the [X/Mg] vs [Mg/H] plane (their Fig.\,17) and that abundance tracks in that space are independent of the stellar sample's spatial position within the Galaxy (their Fig.\,3).

Following \citet{Philcox_2019}, to avoid unrealistic star formation histories (that are very `bursty' for early stars), we additionally require that the SFR (parametrized by a $\Gamma$ distribution with shape parameter $a=2$) at the maximum possible stellar birth-time ($13.8$\,Gyr) should be at least 5\% of the mean SFR, ensuring that there is still a reasonable chance of forming a star at this time-step. This corresponds to the constraint $\log_{10}\left(\mathrm{SFR}_\mathrm{peak}\right)>0.294$. For this reason, a truncated Normal prior will be used for the SFR parameter. Furthermore, we constrain $T_i$ to the interval $[1,13.8]$\,Gyr (assuming an age of the Universe of 13.8\,Gyr), ignoring any stars formed before $1$\,Gyr, which is justified as these are expected to be rare.

\begin{tiny}
\begin{table*}
\begin{minipage}{\textwidth}
\begin{center}
\caption{Free \texttt{Chempy} parameters for each star, with their prior values and Gaussian widths. Stellar birth-times are set for each star individually from a Uniform prior, based on realistic age estimates.}
\begin{tabularx}{\textwidth}{ >{\raggedleft}p{2.2cm}p{6.5cm}|c c }
Parameter & Description & $\overline{\theta}_\mathrm{prior}\pm\sigma_\mathrm{prior}$ & Prior from: \\

\hline
\multicolumn{4}{c}{$\vec{\Lambda}$: \textit{Global stellar (SSP) parameters}}\\
\hline
$\alpha_\mathrm{IMF}$ & High-mass slope of the \citep{2003PASP..115..763C} IMF & $-2.3\pm0.3$ & \citep[Tab.\,1]{2003PASP..115..763C} \\

$\log_{10}\left(N_\mathrm{Ia}\right)$ & Number of SN\,Ia per $\mathrm{M}_\odot$ over 15\,Gyr & $-2.75\pm0.3$ & \citep[Tab.1\,]{2012PASA...29..447M}\\

\hline
\multicolumn{4}{c}{$\vec{\Theta}_i$: \textit{Local ISM parameters}}\\

\hline
$\log_{10}\left(\mathrm{SFE}\right)$ & Star formation efficiency governing gas infall & $-0.3\pm0.3$ & \citep{2008AJ....136.2846B}\\

$\log_{10}\left(\mathrm{SFR}_\mathrm{peak}\right)$ & SFR peak in Gyr (scale of $k=2$ $\Gamma$-distribution) & $0.55\pm0.1$ & \citep[fig.\,4b]{2013ApJ...771L..35V}\\

x$_\mathrm{out}$ & Stellar feedback fraction & $\phantom{-}0.5\pm0.1$ & \citep[Tab.\,1]{Rybizki_2017}\\

\hline
\multicolumn{4}{c}{$T_i$: \textit{Timescale}}\\

\hline
$T_i$ & Time of stellar birth in Gyr & [$1$,$13.8$] & Observations

\label{tab:priors}
\end{tabularx}
\end{center}
\end{minipage}
\end{table*}
\end{tiny}

\paragraph{Nucleosynthetic yield tables}
We adopt the same nucleosynthetic yield tables as in \citep{Philcox_2019}, see their Sec.~2.2 for more details.
To test our method, we aim further at inferring parameters from a sample of stars taken from a hydrodynamical simulation of a MW type galaxy which we take from the Ilustris TNG project \citep{Pillepich2018}. To ensure maximal compatibility with TNG, we adopt their nucleosynthetic yield tables in \texttt{Chempy}, for enrichment by SN\,Ia, SN\,II and AGB stars. The utilized yields are summarized in Tab.\,\ref{tab:chempy_TNG_yields}, matching \citet[Tab.\,2]{2018MNRAS.473.4077P}, and we note that the SN\,II yields are renormalized such that the IMF-weighted yield ratios at each metallicity are equal to those from the \citet{2006ApJ...653.1145K} mass range models alone. \texttt{Chempy} uses only net yields, such that they provide only newly synthesized material, with the remainder coming from the initial SSP composition. These tables may not well-represent true stellar chemistry, and the effects of this missmatch are examined in Sec.\,\ref{subsec: mocks_wrong_yield} by performing inference using an alternative set of yields that does not match the yield set of the training data. For the analysis of observational data, we would want to use the most up-to-date yields, such as \citet{2016ApJ...825...26K} AGB yields, and carefully chose elements which are known to be well reproduced by our current models (e.g. shown by \citet{2019ApJ...874..102W,2019arXiv190806113G}), though this is not appropriate in our context. To facilitate best comparison with TNG, we further set the maximum SN\,II mass as $100\,\mathrm{M}_\odot$ (matching the IMF upper mass limit), adopt stellar lifetimes from \citet{1998A&A...334..505P} and do not allow for any `hypernovae' \citep[in contrary to][]{2018ApJ...861...40P}).

\begin{table}[]
\caption{Nucleosynthetic yield tables used in this analysis, matching those of the TNG simulation \citep[Tab.\,2]{2018MNRAS.473.4077P}.}
     \centering
     \begin{tabular}{c|c}
       Type & Yield Table \\
        \hline
         SN\,Ia & \citet{1997NuPhA.621..467N}\\
         SN\,II & \citet{2006ApJ...653.1145K,portinari}\\
         AGB & \citet{2010MNRAS.403.1413K,2014MNRAS.437..195D};\\
         & \citet{2014ApJ...797...44F}
     \end{tabular}
 \label{tab:chempy_TNG_yields}
 \end{table}

\paragraph{Chemical elements}
In our analysis we only track nine elements: C, Fe, H, He, Mg, N, Ne, O and Si since these are the only elements traced by TNG. We principally compare the logarithmic abundances [X/Fe] and [Fe/H] defined by 
\begin{equation}
     [\mathrm{X}/\mathrm{Y}] = \log_{10}(N_\mathrm X/N_\mathrm Y)_\mathrm{star} - \log_{10}(N_\mathrm X/N_\mathrm Y)_\odot
\end{equation}
for number fraction $N_\mathrm X$ of element X. Here $\odot$ denotes the solar number fractions of \citet{2009ARA&A..47..481A}. This uses H for normalization, thus we are left with $n_\mathrm{el}=8$ independent elements which must be tracked by \texttt{Chempy}\footnote{In observational contexts, it may be more appropriate to compute abundances relative to Mg rather than Fe (as in \citep{2019ApJ...874..102W}) since Mg is only significantly produced by SN\,II and hence a simpler tracer of chemical enrichment.}.

With these modifications, \texttt{Chempy} allows for fast prediction of TNG-like chemical abundances for a given set of galactic parameters. It is important to note that the two GCE models have very different parametrizations of galactic physics, with TNG including vastly more effects, thus it is not certain \textit{a priori} how useful \texttt{Chempy} will be in emulating the TNG simulation, although its utility was partially demonstrated in \citet{2018ApJ...861...40P}. However,  such a test is necessary to prepare for an inference on real data.

\subsection{Neural network emulator for \texttt{CHEMPY}:}
Despite the simplifications made by the GCE model \texttt{CHEMPY}, the run-time of \texttt{CHEMPY} and the high-dimensionality of the parameter space incurs some difficulties when sampling the distribution of the global parameters $\vec\Lambda = \{\alpha_\mathrm{IMF},\log_{10}(N_\mathrm{Ia})\}$. 
To alleviate this, we follow \cite{Philcox_2019} and implement a \textit{neural network} (NN) emulator of the \texttt{CHEMPY} simulator. We design the NN as a simple feed-forward neural network with 2 hidden layers and 100 neurons in the first and 40 neurons in the second layer. Training this tiny emulator takes about 50s on the CPU.

In essence, instead of computing the full model for each input parameter set, we pass the parameters to the NN which predicts the output abundances to high accuracy. As already argued in \cite{Philcox_2019} this has two benefits;
\begin{enumerate}
    \item \textbf{Speed:} The run-time of the \texttt{Chempy} function is $\sim1$\,s per input parameter set, which leads to very slow generation of training data for SBI. With the NN emulator, this reduces to $\sim5\times10^{-5}$\,s, and is trivially parallelizable, unlike \texttt{Chempy}.
    \item \textbf{Differentiability:} The NN is written in pytorch which allows for automatic differentiation. Additionally the NN has a simple closed-form analytic structure \citep[described in the appendix of][]{Philcox_2019}, unlike the complex \texttt{Chempy} model. This allows it to be differentiated, so one can use it to sample via advanced methods such HMC as done in \citep{Philcox_2019}.
\end{enumerate}

Despite the additional complexity introduced by using multiple stellar data-points, our NN simply needs to predict the birth-time abundances for a single star (with index $i$) given a set of six parameters; $\{\vec\Lambda,\vec\Theta_i,T_i\}$. The same NN can be used for all $n_\mathrm{stars}$ stars (and run in parallel), reducing a set of $n_\mathrm{stars}$ runs of \texttt{Chempy} to a single matrix computation. With the above network parameter choices, the NN predicts abundances with an absolute percantage error of \variable{output/ape_NN.txt}, which is far below typical observational errors and even smaller away from the extremes of parameter space (see Fig.~\ref{fig:ape_NN}).
Our NN emulator is publicly available on the github repository accompanying this manuscript \href{https://github.com/TobiBu/sbi-chempy/blob/main/src/scripts/train_torch_chempy.py}{https://github.com/TobiBu/sbi-chempy/blob/main/src/scripts/train\_torch\_chempy.py} with pre-trained NN weights available on zenodo \href{https://zenodo.org/records/14507134}{https://zenodo.org/records/14507134}.

\begin{figure}[]
     \centering
     \includegraphics[width=\columnwidth]{figures/ape_NN.pdf}
     \vspace{-.5cm}
     \caption{Absolute percentage error of the NN emulator for the \texttt{CHEMPY} simulator. The orange histogram shows the full distribution of percentage errors with the vertical dashed line indicating the median and the vertical dotted lines indicating the first and third quartile. The box plot on the top of the plot extends from the first quartile to the third quartile of the data, with a line at the median. The whiskers extend from the box to the farthest data point lying within $1.5\times$ the inter-quartile range from the box. The NN predicts abundances with an absolute percentage error far below typical observational errors.}
     \label{fig:ape_NN}
     \script{train_torch_chempy.py}
\end{figure}

% %The larger volume of data should be able to give tighter statistical constraints on those parameters that are held fixed across the galaxy, but complexity is added since we must allow each star to carry its own set of local ISM parameters.

\paragraph{Bayesian model}
\texttt{CHEMPY} effectively is a Bayesian model for stellar abundances given a set of parameters $\{\Vec{\Lambda},\Vec{\Theta},T\}$ and \citet{Philcox_2019} extended the \texttt{CHEMPY} framework to be able to model multiple stellar data-points. Consider a given star with index $i$ that is born in some region of the ISM. This star will carry its own set of parameters $\{\Vec{\Lambda},\Vec{\Theta}_i,T_i\}$, where $\Vec{\Lambda}$ are star independent and hence taken to be global parameters while the ISM parameters $\Vec{\Theta}_i$ and the birth-time $T_i$ are star specific. Using \texttt{CHEMPY} (or the trained neural network emulator) we can easily model the set of $n_\mathrm{el}$ chemical abundances $\{X_i^j\}$ for the $i$-th star as: 
\begin{equation}\label{eq:chempy_function}
\{X_i^j\} = \texttt{CHEMPY}(\Vec{\Lambda},\Vec{\Theta}_i,T_i),
\end{equation}
where $j$ indexes the chemical element. 
These model abundances can then be compared against observations, with measured abundances $d_i^j$ and corresponding Gaussian errors $\sigma_{i,\mathrm{obs}}^j$ jointly denoted as $D_i=\{d_i^j,\sigma_{i,\mathrm{obs}}^j\}$.

As stated in \citet{Philcox_2019} the full posterior for this case is given by 
\begin{eqnarray}\label{eq:posterior}
    \mathbb{P}(\vec\Lambda,\{\vec\Theta_i\},\{T_i\}|\{D_i\}) &\propto&  \left[\prod_{i=1}^{n_\mathrm{star}}p_{\vec\Theta}(\vec\Theta_i)p_{T_i}(T_i)\right]
    \times p_{\vec\Lambda}(\vec\Lambda)\\
    \nonumber
    &\times& \mathcal{L}(\{D_i\}|\vec\Lambda,\{\vec\Theta_i\},\{T_i\})
\end{eqnarray}
where $p_V(V_i)$ is the prior on variable $V_i$ (belonging to the set $V$).

In order to determine the optimal values of the global galactic parameters ($\Vec{\Lambda}$) one has to sample the posterior of Eq.~\ref{eq:posterior}. This is in practices a fairly costly computation and even in the case of advanced techniques such as Hamilton Monte Carlo sampling the posterior can only be evaluated for a small set of stars ($\lesssim200$) using fairly long compute times \citep[$\sim40$ hours][]{Philcox_2019}.
%
However, recent advances in implicit-likelihood inference or simulation-based inference \citep{Cranmer2020} offer another very efficient approach to approximate the posterior (see next paragraph for more details). These methods train a neural conditional density estimator to represent the conditional posterior, $\mathbb{P}(\vec\Lambda|\{D_i\})$, which can be very efficiently evaluated given observational data $\{D_i\}$.

In particular, if we marginalize over the star specific parameters and solely focus on the global parameters $\Vec{\Lambda}$ we can make the assumption that individual observations of stars are identically and independently distributed (i.i.d.) and factorize the joint posterior from above to simply express it as:  %
\begin{eqnarray}
\mathbb{P}(\vec\Lambda|\{D_i\}) &\propto& 
\mathbb{P}(\vec\Lambda)\mathbb{P}(\{D_i\}|\Lambda)\quad\quad\quad\quad\quad\quad\,\, \text{(Bayes rule)}
\\\nonumber
&=&
\mathbb{P}(\vec\Lambda)\mathbb{P}
(D_1,...,D_{n_\mathrm{star}}|\vec\Lambda)
\\\nonumber
&=& \mathbb{P}(\vec\Lambda)\prod_{j=1}^{n_\mathrm{star}}\mathbb{P}(D_j|\vec\Lambda) 
\\\nonumber
&\propto& \mathbb{P}(\vec\Lambda) \prod_{j=1}^{n_\mathrm{star}}\frac{\mathbb{P}(\vec\Lambda|D_j)}{\mathbb{P}(\vec\Lambda)} \quad\quad\quad\quad\quad\text{(Bayes rule)}
\\%\nonumber
\label{eq:posterior}
&=& \mathbb{P}(\vec\Lambda)^{1-n_\mathrm{star}}\ \prod_{j=1}^{n_\mathrm{star}}\mathbb{P}(\vec\Lambda|D_j) 
\end{eqnarray}

This factorization of the joint posterior into single star posteriors $\mathbb{P}(\vec\Lambda|D_j)$ has the advantage that we do not need to specify the exact number of observational data points before. We can simply train our neural density estimator on single star observations and then at evaluation time, we simply combine as many posterior estimates as we have observational data.

\paragraph{Simulation-based inference}
In a nutshell, SBI \citep[e.g.][]{Cranmer2020,Papamakarios:2021,Gloeckler2024AllinoneSI} -- also called likelihood free inference within a Bayesian inference framework -- works as follows: given an assumed generative model $\mathcal{M}$ of parameters $\Vec{\theta}$ (in our case a galactic chemical evolution model) and a set of simulated observations of individual stellar abundances $\Vec{X}$ from that model, we train a mapping between the two to estimate the posterior distribution $p$($\Vec{\theta}|\Vec{X}$, $\mathcal{M}$) of the generative model parameters $\Vec{\theta}$ that reproduce the simulated observations $\Vec{X}$. Once this mapping is trained, we can apply it to observations of stellar abundances $\Vec{X_R}$ to infer $p$($\Vec{\theta}|\Vec{X_R}$, $\mathcal{M}$) (see Fig. \ref{fig:flowchart} for a schematic visual representation of our method). Note, we do not need to know anything about $p$($\Vec{\theta}|\Vec{X_R}$, $\mathcal{M}$), we solely need to be able to sample from it.

We use a Neural Posterior Estimator (NPE) \cite{zeghal2022neuralposteriorestimationdifferentiable} which utilises the gradients of the generative model $\mathcal{M}$ with a Masked Autoregressive Flow model (MAF) \cite{papamakarios2018maskedautoregressiveflowdensity} containing 10 hidden features and 1 transformation layer for the normalizing flow.

We train our NPE with $10^5$ data points consisting of $n_{\rm elemets}=8$ simulated with the NN emulator described in the previous section with inputs sampled from the priors shown in Tab. \ref{tab:priors}. Training takes $\lesssim10$ minutes on an Intel(R) Xeon(R) Platinum 8468V with 192 CPUs. 
We evaluate the accuracy of the NPE using $5\times10^4$ validation data points from the original \texttt{CHEMPY} simulator. In order to mimick observational noise, we add 5\% observational uncertainties to the abundances simulated with \texttt{CHEMPY} before feeding them to our NPE.
We have made sure that our NPE is well calibrated and posterior distributions are trustable. For more details on simulation-based calibration see appendix \ref{sec:sbc}.

\begin{figure}[]
     \centering
     \includegraphics[width=\columnwidth]{figures/ape_posterior_NPE_C.pdf}
     \vspace{-.5cm}
     \caption{Absolute percentage error of the neural posterior density estimate for a single star. Different coloured histograms show the full error distribution for all 6 parameters of interest with the median values highlighted by the vertical dashed lines. The box plots show again the first and third quartiles with the median represented by a vertical line and the whiskers extending from the box to the farthest data point lying within $1.5\times$ the inter-quartile range from the box. The global parameters of main interest for this work are shown by the light blue and red histogram.}
     \label{fig:posterior_APE}
     \script{plot_posterior_APE.py}
\end{figure}

\paragraph{Multi-star inference}
Following equation~\ref{eq:posterior} we can calculate the joint posterior for a combined inference using abundance observations of multiple stars. For this, we will condition our NPE individually on single star observations and sample the posterior. Then, according to equation~\ref{eq:posterior} we simply have to multiply individual posteriors and account for the inverse weighting by some power of the prior which we take from table~\ref{tab:priors}.

Note however, there is one issue here. Since we only have access to samples from the posterior and not the posterior itself it is difficult to evaluate equation~\ref{eq:posterior}. We circumvent this by approximating each single star posterior by a Gaussian and fit for the parameters of mean and covariance. With this it is straight forward to evaluate equation~\ref{eq:posterior} analytically. In fact, under our assumption the combined posterior is a product of the prior and the product of multiple Gaussians for the individual star posteriors. The latter product is also a Gaussian with mean $\mathbf{\mu'}$ and variance $\mathbf{\sigma'}$:
\begin{align}
    \mathbf{\mu'} &= \frac{\sum_{i=1}^{N_{stars}} \frac{\mu_i}{\sigma_i^2}}{\sum_{i=1}^{N_{stars}} \frac1{\sigma_i^2}} \\
\mathbf{\sigma'}^2 &= \frac1 {\sum_{i=1}^{N_{stars}} \frac1{\sigma_i^2}}
\end{align}

Further, in our case the prior for the galactic parameters $\Lambda$ is Gaussian as well. Therefore the resulting factorized posterior from eq.~\ref{eq:posterior} is again a Gaussian and can be expressed with mean $\mathbf{\mu}$ and variance $\mathbf{\sigma}$ as:
\begin{align}
\mathbf{\mu} &= \frac{\frac{\mu'}{\sigma'^2}-\frac{(1-N)\mu_ {prior}}{\sigma_ {prior}^2}}{\frac1{\sigma'^2}-\frac{(1-N)}{\sigma_ {prior}^2}} \\
\mathbf{\sigma}^2 &= \frac1 {\frac1{\sigma'^2}-\frac{(1-N)}{\sigma_ {prior}^2}}
\end{align}

Note, given the tiny and simple neural network that represents our NPE the above assumption of Gaussian single star posteriors and the consecutive simplification is a decent one to make. We have further empirically verified that single star posteriors are indeed close to Gaussian.
However, in future work we plan to alleviate this simplification and directly approximate the joint posterior of a multi-star inference.

\section{Results}
\label{sec: Results}

We use our SBI method described in the previous section to infer the global galactic parameters $\alpha_{\rm {IMF}}$ and $\log_{10}(N_{\rm{Ia}})$. In order to demonstrate the performance and robustness of our methods we use three mock data-sets: 
\begin{enumerate}
    \item Mock observations drawn from \texttt{CHEMPY} from the same yield set as the training data for the neural network emulator. With this we esnure to test our training strategy and the performance without any systematic distribution shifts. 
    \item  \texttt{CHEMPY} mock data using a different yield set to test for potential biases through model missspecification in our SBI setup.
    \item Simulated data from a large-scale hydrodynamical simulation taken from the IlustrisTNG suite \citep{Pillepich2018} but with the same yield set as our \texttt{CHEMPY} training data to ensure that we recover the correct parameters even for models with a completely different treatment of ISM physics.
\end{enumerate}

Note, the methods presented here can naturally be extended to any fast and flexible GCE model, not just \texttt{Chempy}.

\subsection{\texttt{CHEMPY} Mock Observational Data}
Our analysis uses mock observations drawn from the \textit{neural network} emulator with fixed values of the global galactic parameters $\alpha_{\rm IMF}=-2.3$ and $\log_{10}(N_{\rm Ia})=-2.89$ and varying local parameters $\vec{\Theta}_i$ using priors from Tab. \ref{tab:priors}. Additionally we draw $T_i$ uniformly in the range $[2,12.8]$ Gyr to minimize overlap with the neural network training birth-time limits when observational uncertainties are included. 
Each set of parameters is passed to our \texttt{CHEMPY} emulator, producing eight true chemical abundances. In order to fully represent observational data, we augment this data with observational uncertainties by adding a Gaussian error of $0.05$ dex for the abundances representative for typical APOGEE data \citep{Majewski2016}. All our datasets and accompanying code to reproduce our data generation process and analysis is publicly available\footnote{\url{https://github.com/TobiBu/sbi-chempy}}.

For each individual observation consisting of the abundance measurements of a single star we sample the posterior estimate for all six parameters $\{\vec\Lambda,\vec\Theta_i,T_i\}$ with 1000 points. This takes around $0.3s$ for each star, making it extremely fast to infer the parameters of a large amount of stars. Our method takes around $10$ seconds to build a posterior for all six parameters for a dataset size of 1000 stars (each time sampling the single star posterior with 1000 points and fitting for the Gaussian parameters). Our combined runtime for the NPE training plus sampling is hence 610 seconds for 1000 stars hence our methods making it more than $1.180$ times faster than current HMC methods \citep[cf.][who needs $40$h for only 200 stars]{Philcox_2019}. But since our approach is amortized and we do not need to retrain our NPE model each time we want to make an inference, we in fact achieve an inference speed-up of a factor of $\sim72.000$. \textcolor{red}{check if those speed-up values are correctly calculated...}
Hence, shorter computing times make it feasible to use orders of magnitudes more observations.

\paragraph{Validation through absolute percentage error}
We start evaluating our method by using \texttt{CHEMPY} to produce a mock observational dataset to ensure no systematical shft between training and testing data in terms of physics parameters.

To evaluate our method quantitatively, we compare the posterior mean to the ground truth value for each observation and calculate the absolute percentage error (APE, see Fig.~\ref{fig:posterior_APE}). For a single star observation, our NPE has an APE of \variable{output/global_posterior_APE.txt} when looking at all 6 parameters of interest. If we restrict ourselves to only the global two parameters $\vec\Lambda = \{\alpha_\mathrm{IMF},\log_{10}(N_\mathrm{Ia})\}$ our NPE achieves an APE of \variable{output/posterior_APE.txt} as we show in Fig.~\ref{fig:posterior_APE}. Looking at Fig.~\ref{fig:posterior_APE} we find that for an individual star the accuracy of the NPE is not particularly high. However, for the global parameters $\vec\Lambda = \{\alpha_\mathrm{IMF},\log_{10}(N_\mathrm{Ia})\}$ we can boost the accuracy by combining the inference for many stars of the same galaxy.

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figures/CHEMPY_TNG_yields_Nstar_comp.pdf}
    \vspace{-.5cm}
    \caption{Accuracy of inferred global galactic parameters $\alpha_{IMF}$ and $\log_{10}(N_{Ia})$ as a function of number of observed stars, comparing SBI (blue line) to the inferred values using HMC (red line) as done by \cite{Philcox_2019} and the ground truth values (black dashed line). For the SBI analysis we show $1\sigma$ and $2\sigma$ contours while HMC results only show $1\sigma$ statistical uncertainities as reported in Tab.~3 of \citet{Philcox_2019} (blue/red shaded regions).}
    \label{fig:CHEMPY_TNG_N_star_analysis}
    \script{chempy_tng_inference.py}
\end{figure*}

\paragraph{Inference on the \texttt{CHEMPY} TNG yield set}

\begin{figure}
%\vspace{-1.3cm}
    \centering
    \includegraphics[width=\columnwidth]{figures/CHEMPY_TNG_yields.png}
    \vspace{-.25cm}
    \caption{Joint posterior for global galactic parameters $\alpha_{\rm IMF}$ and $\log_{10}(\rm N_{Ia})$ of $1000$ stars. The ground truth value is shown by the black dot, the posterior mean of the SBI inference is shown with the red star and white and black ellipses show the $1-3\sigma$ contours of our inference.}
    \label{fig:CHEMPY_TNG_sbi} 
    \script{chempy_tng_inference.py}
\end{figure}

%
Combining the inference on multiple observed stars gives us higher accuracy and precision of the global galactic parameters $\vec\Lambda$. We therefore perform inference using a range of stars $n\in[1,1000]$ (see Fig. \ref{fig:CHEMPY_TNG_N_star_analysis}) to show how inference accuracy increases with number of observations. We find that in the limit of few stars (less than $\sim100$) SBI shows a larger uncertainties than the HMC results of \citet{Philcox_2019}. But in the limit of large numbers of stars (few hundred) the accuracy and precision is superior compared to HMC. In particular, after using a few ten stars, the NPE estimate is less biased than the HMC results. 
Finally, given our computational advantage we will be able to use orders of magnitude more stars for our inference.
This is particular important since sample variances play an important role when using small samples of stars as also noted in \citet{Philcox_2019}.
%

In Fig.~\ref{fig:CHEMPY_TNG_sbi} we show the joint posterior for $\alpha_{IMF}$ and $\log_{10}(N_{Ia})$ for our inference. The red star indicates the ground truth value, the black dot shows our posterior mean and white contours indicate $1-3\sigma levels$.

Using a sample of 1000 stars we infer \variable{output/CHEMPY_TNG_sbi.txt} which deviates less than $\sim0.04\%$ from the correct value.

Note, as shown in Fig.~\ref{fig:posterior_APE} in principle our analysis is also able to infer the local parameters $\Vec{\Theta}_i$ and $T_i$. This would allow us to estimate/infer stellar ages as well. However, currently our NPE network is not particularly good at estimating ages from abundances alone. On avergae our inference for ages results in an APE of $\sim27\%$ which is slightly larger than the observational noise of $20\%$ that we add during mock up of our data.

In summary, our SBI pipeline is quite capable of correctly and precisely inferring  global parameters of chemical enrichment models from stellar abundance alone when using the same physical model and yield tables as during training. Next, we will check what happens if for some reason the training data was generated with a different yield set.

\subsection{Alternative yield set}
\label{subsec: mocks_wrong_yield}

\section{Discussion}
\label{sec: conclusion}

\textcolor{red}{here goes the assumption that stars are iid since each stars is produced by a single call to chempy. However, real physics might correlate stars abundacnes by star formation histories and cumulative enrichment...}

\section{Summary and conclusions}
\label{sec: conclusion}

Here we have explored innovative SBI methods to constrain important galactic parameters from the multi-dimensional observation of chemical abundances of stars. Our results show that SBI provides a super fast method, that beats previous inference algorithms by orders of magnitude, and delivers accurate and precise results. Hence, SBI enables faster and more accurate evaluation of observational data after a small initial computational investment of training a NPE ($\sim3$ h). Since evaluation of the NPE is only weakly dependent on the dimensionality of the input observations we expect almost perfect scaling to larger chemical abundances observations and even further improvements in the inference results with future high quality data from massive spectroscopic surveys.

\begin{acknowledgements}
      This project was made possible by funding from the Carl Zeiss Stiftung.
\end{acknowledgements}

% WARNING
%-------------------------------------------------------------------
% Please note that we have included the references to the file aa.dem in
% order to compile it, but we ask you to:
%
% - use BibTeX with the regular commands:
%   \bibliographystyle{aa} % style aa.bst
%   \bibliography{Yourfile} % your references Yourfile.bib
%
% - join the .bib files when you upload your source files
%-------------------------------------------------------------------

\bibliographystyle{aa}
\bibliography{bib.bib}

\begin{appendix}

\section{Code and Data Availability}
\label{sec:appendix_code_and_data}

To facilitate a wider community's usage and contributions, we make use of the reproducibility software
\href{https://github.com/showyourwork/showyourwork}{\showyourwork}
\citep{Luger2021}, which leverages continuous integration to
programmatically download the data from
\href{https://zenodo.org/}{zenodo.org}, create the figures, and
compile the manuscript. Each figure caption contains two links: one
to the dataset stored on zenodo used in the corresponding figure,
and the other to the script used to make the figure (at the commit
corresponding to the current build of the manuscript). The git
repository associated to this study is publicly available at
\url{https://github.com/TobiBu/sbi-chempy}, and the release
v0.4.1 allows anyone to re-build the entire manuscript including rerunning all analysis. The datasets and neural network weights are stored at \url{https://doi.org/10.5281/zenodo.7343715} \textcolor{red}{change to final zenodo dataset!}.

% The ... repository, which contains code, additional documentation, and interactive dashboards, is available on GitHub.%\footnote{URL: \url{https://github.com/bGuenes/sbi_chemical_abundances}}

\section{Simulation-based calibration}
\label{sec:sbc}

\begin{figure*}[]
     \centering
     \includegraphics[width=\textwidth]{figures/sbc_rank_plot_NPE_C.pdf}
     %\vspace{-.5cm}
     \caption{SBC ranks of ground truth parameters under the inferred posterior samples for each of the six parameters (red bars). The grey area shows the 99\% confidence interval of a uniform distribution given the number of samples provided.}
     \label{fig:sbc}
     \script{evaluate_sbi.py}
\end{figure*}

\begin{figure}[]
     \centering
     \includegraphics[width=\columnwidth]{figures/tarp_plot_NPE_C.pdf}
     %\vspace{-.5cm}
     \caption{TARP plot showing the expected coverage probability vs. the credibility level $\alpha$. The dashed black 1:1-line shows an ideal calibrated posterior and the blue solid line shows the TARP value for our NPE.}
     \label{fig:tarp}
     \script{evaluate_sbi.py}
\end{figure}

Since SBI relies in neural networks to approximate posterior densities one important point is to check that neural network hyperparameters are well chosen and that posterior estimates are trustable.

After a density estimator has been trained with simulated data to obtain a posterior, the estimator should be made subject to several diagnostic tests. This needs to be performed before being used for inference given the actual observed data. Posterior Predictive Checks provide one way to "critique" a trained estimator based on its predictive performance. Another important approach to such diagnostics is simulation-based calibration as developed by \citet{Cook2006} and \citet{Talts2018}.

\paragraph{Simulation-based calibration}
Simulation-based calibration (SBC) provides a (qualitative) view and a quantitive measure to check, whether the variances of the posterior are balanced, i.e. it is neither over-confident nor under-confident. As such, SBC can be viewed as a necessary condition (but not sufficient) for a valid inference algorithm: If SBC checks fail, this tells you that your inference is invalid. If SBC checks pass, this is no guarantee that the posterior estimation is working.

To perform SBC, we sample some $\theta_i^o$ values from the parameter prior of the problem at hand and simulate "observations" from these parameters: 
\begin{equation}
    x_i = \text{simulator}(\theta_i^o)
\end{equation}
Then we perform inference given each observation $x_i$ which produces a separate posterior $p_i(\theta|x_i)$ for each $x_i$. The key step for SBC is to generate a set of posterior samples $\{\theta\}_i$ from each posterior. We call this $\theta_i^s$, referring to $s$ samples from the posterior $p_i(\theta|x_i)$. Next, we rank the corresponding $\theta_i^o$ under this set of samples. A rank is computed by counting how many samples $\theta_i^s$ fall below their corresponding $\theta_i^o$ value \citep[see section 4.1 in][]{Talts2018}. These ranks are then used to perform the SBC check itself.

The core idea behind SBC is two fold: (i) SBC ranks of ground truth parameters under the inferred posterior samples follow a uniform distribution (If the SBC ranks are not uniformly distributed, the posterior is not well calibrated); and (ii) samples from the data averaged posterior (ensemble of randomly chosen posterior samples given multiple distinct observations $x_o$) are distributed according to the prior.

Hence, SBC can tell us whether the SBI method applied to the problem at hand produces posteriors that have well-calibrated uncertainties, and if the posteriors have uncalibrated uncertainties, SBC surfaces what kind of systematic bias is present: negative or positive bias (shift in the mean of the predictions) or over- or under-dispersion (too large or too small variance).

In the left panel of Fig.~\ref{fig:sbc} we show the distribution of ranks (depicted in red) in each dimension. Highlighted in grey, you see the 99\% confidence interval of a uniform distribution given the number of samples provided. In plain english: for a uniform distribution, we would expect 1 out of 100 (red) bars to lie outside the grey area. This figure shows that overall our posteriors are decently calibrated. Only for the parameter $\log_{10}\left(\mathrm{SFE}\right)$ we see a single bar outside the 99\% confidence interval. But most importantly for the parameters of interest here, $\log_{10}\left(N_\mathrm{Ia}\right)$ and $\alpha_\mathrm{IMF}$ we have a well calibrated posterior.

\paragraph{Tests of Accuracy with Random Points (TARP)}

TARP \citep{Lemos2023} is an alternative calibration check. Given a test set $(\theta^*,x^*)$ and a set of reference points $\theta_r$, TARP calculates statistics for posterior calibration by - drawing posterior samples $\theta$ given each observation $x^*$ and calculating the distance $r$ between $\theta^*$ and $\theta_r$ counting for how many of the posterior samples the distance to $\theta_r$ is smaller than $r$ \citep[see e.g. Fig.~2 in][for an illustration]{Lemos2023}.

For each given coverage level $\alpha$, one can then calculate the corresponding average counts and check, whether they correspond to the given $\alpha$. The visualization and interpretation of TARP values is therefore similar to that of SBC. However, in contrast to SBC, TARP provides a necessary and sufficient condition for posterior accuracy, i.e., it can also detect inaccurate posterior estimators. In Fig.~\ref{fig:tarp} we show the result for our NPE in blue in comparison to the ideal line shown in black dashed style. This figure clearly shows that our NPE is well calibrated.

Note, however, that this property depends on the choice of reference point distribution: to obtain the full diagnostic power of TARP, one would need to sample reference points from a distribution that depends on $x$. Thus, in general, it is recommended using and interpreting TARP like SBC and complementing coverage checks with posterior predictive checks.

\end{appendix}

\end{document}

%%%% End of aa.dem